---
title: "APTC Auto Re-Enrollment Elimination Analysis Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: flatly
    navbar:
      - { title: "GitHub", href: "https://github.com/YOUR_USERNAME/YOUR_REPO", align: right, icon: "fa-github" }
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(leaflet)
library(plotly)
library(scales)
library(sf)
library(tigris)
library(htmltools)

options(tigris_use_cache = TRUE)

df <- read_csv("~/Downloads/aptc_clean - aptc_clean.csv.csv")

nat_current       <- sum(df$current_uncomp_spending,       na.rm = TRUE)
nat_post          <- sum(df$post_policy_uncomp_spending,    na.rm = TRUE)
nat_change        <- sum(df$change_in_uncomp_spending,      na.rm = TRUE)
nat_lost          <- sum(df$post_policy_aptc_lost_coverage, na.rm = TRUE)
nat_uninsured_new <- sum(df$post_policy_aptc_uninsured,     na.rm = TRUE)

us_pop           <- sum(df$state_population, na.rm = TRUE)
nat_current_rate <- sum(df$current_uninsured_rate    * df$state_population, na.rm = TRUE) / us_pop
nat_post_rate    <- sum(df$post_policy_uninsured_rate * df$state_population, na.rm = TRUE) / us_pop

states_sf <- states(cb = TRUE, resolution = "20m", year = 2023) %>%
  filter(!STUSPS %in% c("PR", "VI", "GU", "MP", "AS", "UM")) %>%
  st_transform(crs = 4326)

map_data <- states_sf %>%
  left_join(df, by = c("STUSPS" = "state"))

map_colors    <- c("#BEE7B8", "#8DA267", "#5E663D", "#7B2D42")
legend_breaks <- c(0, 50e6, 200e6, 600e6, 1300e6)
legend_labels <- c("< $50M", "$50M – $200M", "$200M – $600M", "> $600M")
legend_colors <- c("#BEE7B8", "#8DA267", "#5E663D", "#7B2D42")

pal_map <- colorBin(
  palette  = map_colors,
  domain   = map_data$change_in_uncomp_spending,
  bins     = legend_breaks,
  na.color = "#cccccc"
)

col_current  <- "#5E663D"
col_increase <- "#7B2D42"

fmt_uncomp <- function(x) {
  ifelse(abs(x) >= 1e9,
         paste0("$", round(x / 1e9, 2), "B"),
         paste0("$", round(x / 1e6, 1), "M"))
}

popup_html <- paste0(
  "<b>", map_data$NAME, " (", map_data$STUSPS, ")</b><br/>",
  "<hr style='margin:4px 0;'>",
  "<b>Current Uncomp. Care:</b> ",     fmt_uncomp(map_data$current_uncomp_spending),    "<br/>",
  "<b>Post-Policy Uncomp. Care:</b> ", fmt_uncomp(map_data$post_policy_uncomp_spending), "<br/>",
  "<b>Increase in Costs:</b> ",        fmt_uncomp(map_data$change_in_uncomp_spending),  "<br/>",
  "<b>Projected New Uninsured:</b> ",
  formatC(map_data$post_policy_aptc_uninsured,     format = "d", big.mark = ","), "<br/>",
  "<b>Coverage Lost:</b> ",
  formatC(map_data$post_policy_aptc_lost_coverage, format = "d", big.mark = ","), "<br/>",
  "<b>Current Uninsured Rate:</b> ",
  round(map_data$current_uninsured_rate    * 100, 1), "%<br/>",
  "<b>Post-Policy Uninsured Rate:</b> ",
  round(map_data$post_policy_uninsured_rate * 100, 1), "%"
)
```

<style>
.navbar-nav > li > a > .fa { display: none !important; }
.navbar { background-color: #6B3A4A !important; border-color: #7B2D42 !important; }
.navbar .navbar-brand, .navbar-nav > li > a { color: #F5EEF0 !important; }
.navbar-nav > li.active > a, .navbar-nav > li > a:hover {
  color: #ffffff !important; background-color: #7B2D42 !important;
}
.chart-title {
  font-size: 1rem !important; color: #333 !important;
  font-weight: 600 !important; background-color: #f7f7f7 !important;
  border-bottom: 1px solid #e5e5e5 !important;
}
html, body { height: 100%; background-color: #ffffff !important; }
.tab-content, .tab-pane, .flexdashboard-page { background-color: #ffffff !important; }
.dashboard-page-wrapper { min-height: 100vh; background-color: #ffffff !important; }
.section.level2 { background-color: #ffffff !important; }
.tab-pane:not(.active) { display: none !important; }
.tab-pane:not(.active) .plotly { visibility: hidden !important; }
</style>

Home
=====================================

Row
-------------------------------------

###

```{r}
HTML('
<div style="max-width:1200px; margin:40px auto; padding:0 60px;
            font-family:\"Source Sans Pro\", sans-serif; color:#1D1D20; line-height:1.85;">

  <h2 style="font-size:2.4rem; font-weight:700; margin-bottom:16px; color:#300204;">
    What Are APTC Subsidies and What Is Being Proposed About Them?
  </h2>

  <p style="font-size:1.4rem; margin-bottom:18px;">
    <strong>Advanced Premium Tax Credits (APTCs)</strong> are federally subsidized
    payments that reduce monthly health insurance premiums for individuals and families
    purchasing coverage through the ACA Marketplace. Recipients are generally lower-income
    adults who earn between 100% and 400% of the federal poverty level. In 2025, approximately
    14.5 million Americans received APTCs, and roughly half are automatically re-enrolled each year if they do
    not actively change or drop their plan.
  </p>

  <p style="font-size:1.4rem; margin-bottom:18px;">
    The One Big Beautiful Bill Act (OBBBA), currently advancing in Congress, proposes to
    eliminate automatic re-enrollment for APTC recipients. Under this change, every enrollee
    would be required to actively re-confirm their coverage each year. Individuals who fail
    to do so would lose their subsidized coverage at renewal.
  </p>

  <hr style="border:none; border-top:1px solid #ddd; margin:28px 0;">

  <h3 style="font-size:2.4rem; font-weight:700; margin-bottom:10px; color:#300204;">
    Research Question
  </h3>
  <p style="font-size:1.4rem; margin-bottom:18px;">
    If automatic re-enrollment for APTC subsidies is eliminated, how many enrollees are
    projected to lose coverage and what would be the impact on uncompensated care costs
    for U.S. hospitals and safety-net providers?
  </p>

  <hr style="border:none; border-top:1px solid #ddd; margin:28px 0;">

  <h3 style="font-size:2.4rem; font-weight:700; margin-bottom:10px; color:#300204;">
    Data &amp; Methods
  </h3>
  <p style="font-size:1.4rem; margin-bottom:18px; color:#444;">
    Enrollment data are drawn from the CMS 2025 Marketplace Open Enrollment Period Public
    Use Files. Our analysis is based on uninsured rates from the Kaiser Family Foundation
    and operates under the assumption that there will be a 33% reduction in re-enrollment
    following elimination of automatic re-enrollment (Shepard &amp; Wagner, 2025). Of those
    projected to lose coverage, we assume that 50% of those enrollees will become uninsured
    (McIntyre et al., 2024). Uncompensated care costs are estimated at $1,230.73 per newly
    uninsured person per year (Garthwaite et al., 2015) and adjusted to 2026 dollars using
    the U.S. Bureau of Labor Statistics Inflation Calculator (n&nbsp;=&nbsp;51).
  </p>

  <hr style="border:none; border-top:1px solid #ddd; margin:28px 0;">

  <h3 style="font-size:2.4rem; font-weight:700; margin-bottom:10px; color:#300204;">
    Impact Statement: Why This Proposed Change Matters
  </h3>
  <p style="font-size:1.4rem; margin-bottom:18px;">
    While Section 112201 of the One Big Beautiful Bill Act projects $5.45 billion in
    federal subsidy savings, these gains are more than offset by a projected $6.10 billion
    increase in annual uncompensated care costs (CMS, 2025). Eliminating automatic
    re-enrollment introduces an administrative burden that disproportionately harms
    low-income enrollees and vulnerable populations, ultimately reversing coverage gains
    achieved under the ACA and destabilizing the broader insurance market through weakened
    risk pools and higher premiums for all consumers.
  </p>

  <hr style="border:none; border-top:1px solid #ddd; margin:28px 0;">

  <h3 style="font-size:2.4rem; font-weight:700; margin-bottom:10px; color:#300204;">
    About This Dashboard
  </h3>
  <p style="font-size:1.4rem; margin-bottom:18px;">
    This dashboard is designed as an interactive policy tool for researchers, lawmakers,
    advocates, and health system administrators seeking to understand the projected
    state-level and national impacts of eliminating automatic re-enrollment for APTC
    recipients. By allowing users to explore projected uninsured rates, uncompensated care
    costs, coverage loss figures and model how costs shift under alternative uninsured rate
    assumptions, this dashboard translates complex administrative and fiscal projections
    into accessible, actionable insights for evidence-based health policy decision making.
  </p>

  <hr style="border:none; border-top:1px solid #ddd; margin:28px 0;">

  <h3 style="font-size:2.4rem; font-weight:700; margin-bottom:10px; color:#300204;">
    Sources
  </h3>
  <p style="font-size:0.92rem; color:#555; line-height:1.9;">
    Centers for Medicare &amp; Medicaid Services. (2025). <em>2025 Marketplace Open Enrollment Period Public Use Files.</em> U.S. Department of Health and Human Services. https://www.cms.gov/data-research/statistics-trends-reports/marketplace-products/2025-marketplace-open-enrollment-period-public-use-files
    <br/><br/>
    Garthwaite, C., Gross, T., &amp; Notowidigdo, M. J. (2015). <em>Hospitals as insurers of last resort</em> (NBER Working Paper No. 21290). National Bureau of Economic Research. https://www.nber.org/papers/w21290
    <br/><br/>
    Kaiser Family Foundation. (2024). <em>State health facts: Total population by insurance status</em> [Data set]. KFF. https://www.kff.org/state-health-policy-data/state-indicator/total-population/
    <br/><br/>
    McIntyre, A., Sommers, B. D., Aboulafia, G., Phelan, J., Orav, E. J., Epstein, A. M., &amp; Figueroa, J. F. (2024). Coverage and access changes during Medicaid unwinding. <em>JAMA Health Forum, 5</em>(6), e242193. https://doi.org/10.1001/jamahealthforum.2024.2193
    <br/><br/>
    Shepard, M., &amp; Wagner, M. (2025). Do ordeals work for selection markets? Evidence from health insurance auto-enrollment. <em>American Economic Review, 115</em>(3), 772–822. https://doi.org/10.1257/aer.20231133
    <br/><br/>
    U.S. Bureau of Labor Statistics. (2026). <em>CPI inflation calculator.</em> U.S. Department of Labor. https://www.bls.gov/data/inflation_calculator.htm
  </p>

</div>
')
```


About the Data
=====================================

Row
-------------------------------------

###

```{r}
HTML('
<div style="max-width:1200px; margin:40px auto; padding:0 60px;
            font-family:\"Source Sans Pro\", sans-serif; color:#1D1D20; line-height:1.85;">

  <h2 style="font-size:2.4rem; font-weight:700; margin-bottom:20px; color:#300204;">
    Dataset Description
  </h2>

  <p style="font-size:1.3rem; margin-bottom:16px;">
    Data are sourced from the <strong>Centers for Medicare &amp; Medicaid Services (CMS)
    Open Enrollment Public Use Files</strong> for plan year 2025, available at
    <a href="https://data.healthcare.gov" target="_blank">data.healthcare.gov</a>.
  </p>

  <p style="font-size:1.3rem; margin-bottom:16px;">
    The dataset covers all 50 U.S. states and the District of Columbia
    (n&nbsp;=&nbsp;51 state-level observations) and reflects ACA Marketplace enrollment
    activity during the 2025 open enrollment period.
  </p>

  <p style="font-size:1.3rem; margin-bottom:16px;">
    Data were compiled from administrative enrollment records reported by CMS and
    supplemented with author calculations using a per-person uncompensated care cost
    estimate of $1,230.73, derived from Dranove et al. (2015), NBER Working Paper
    No.&nbsp;21290.
  </p>

  <p style="font-size:1.3rem; margin-bottom:16px;">
    The study population consists of ACA Marketplace enrollees receiving Advanced Premium
    Tax Credits (APTCs), with a specific focus on those enrolled through automatic
    re-enrollment across all states for plan year 2025.
  </p>

  <p style="font-size:1.3rem; color:#444;">
    Post-policy projections simulate the elimination of automatic re-enrollment for APTC
    recipients as proposed under the One Big Beautiful Bill Act (OBBBA); it is assumed
    that 50% of individuals who lose APTC-subsidized coverage transition to uninsured
    status.
  </p>

</div>
')
```

Row
-------------------------------------

### State-Level Data

```{r}
library(DT)

df_table <- df %>%
  select(
    State                        = state,
    `Total Consumers`            = total_consumers,
    `APTC Recipients`            = aptc_consumers,
    `% Receiving APTC`           = pct_aptc,
    `Auto Re-enrollees`          = auto_reenrollees,
    `% Auto Re-enrolled`         = pct_auto_reenroll,
    `Coverage Lost`              = post_policy_aptc_lost_coverage,
    `Projected New Uninsured`    = post_policy_aptc_uninsured,
    `Current Uninsured Rate`     = current_uninsured_rate,
    `Post-Policy Uninsured Rate` = post_policy_uninsured_rate,
    `Current Uncomp. Care`       = current_uncomp_spending,
    `Post-Policy Uncomp. Care`   = post_policy_uncomp_spending,
    `Increase in Uncomp. Care`   = change_in_uncomp_spending
  ) %>%
  mutate(
    `% Receiving APTC`           = paste0(round(`% Receiving APTC` * 100, 1), "%"),
    `% Auto Re-enrolled`         = paste0(round(`% Auto Re-enrolled` * 100, 1), "%"),
    `Current Uninsured Rate`     = paste0(round(`Current Uninsured Rate` * 100, 1), "%"),
    `Post-Policy Uninsured Rate` = paste0(round(`Post-Policy Uninsured Rate` * 100, 1), "%"),
    `Current Uncomp. Care`       = dollar(`Current Uncomp. Care`, scale = 1e-6, suffix = "M", accuracy = 0.1),
    `Post-Policy Uncomp. Care`   = dollar(`Post-Policy Uncomp. Care`, scale = 1e-6, suffix = "M", accuracy = 0.1),
    `Increase in Uncomp. Care`   = dollar(`Increase in Uncomp. Care`, scale = 1e-6, suffix = "M", accuracy = 0.1)
  )

datatable(
  df_table,
  rownames   = FALSE,
  filter     = "none",
  extensions = "Buttons",
  options    = list(
    pageLength = 15,
    dom        = "Bfrtip",
    buttons    = list("csv"),
    scrollX    = TRUE,
    columnDefs = list(list(className = "dt-center", targets = "_all"))
  )
)
```


National Level Map
=====================================

Row {data-height=90}
-------------------------------------

###

```{r}
HTML(paste0('
<div style="display:flex; gap:32px; padding:14px 24px; flex-wrap:wrap;
            font-family:\"Source Sans Pro\", sans-serif; color:#1D1D20;">

  <div style="flex:1; min-width:200px;">
    <div style="font-size:1.6rem; font-weight:700; color:#1D1D20;">$',
      round(nat_change / 1e9, 2), 'B</div>
    <div style="font-size:1rem; color:#444;">
      Projected National Increase in Uncompensated Care Costs
    </div>
  </div>

  <div style="flex:1; min-width:200px;">
    <div style="font-size:1.6rem; font-weight:700; color:#1D1D20;">',
      formatC(nat_uninsured_new, format="d", big.mark=","), '</div>
    <div style="font-size:1rem; color:#444;">
      Projected New Uninsured Americans
    </div>
  </div>

  <div style="flex:1; min-width:200px;">
    <div style="font-size:1.6rem; font-weight:700; color:#1D1D20;">',
      formatC(nat_lost, format="d", big.mark=","), '</div>
    <div style="font-size:1rem; color:#444;">
      APTC Recipients Projected to Lose Coverage
    </div>
  </div>

</div>
'))
```

Row {data-height=520}
-------------------------------------

### Projected Increase in Uncompensated Care by State — Click a State for Details

```{r}
leaflet(map_data,
        options = leafletOptions(minZoom = 3, maxZoom = 8)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addPolygons(
    fillColor    = ~pal_map(change_in_uncomp_spending),
    fillOpacity  = 0.85,
    color        = "white",
    weight       = 1,
    smoothFactor = 0.5,
    popup        = popup_html,
    highlight    = highlightOptions(
      weight       = 2.5,
      color        = "#333",
      fillOpacity  = 0.95,
      bringToFront = TRUE
    ),
    label = ~paste0(STUSPS, ": ", fmt_uncomp(change_in_uncomp_spending)),
    labelOptions = labelOptions(
      style     = list("font-weight" = "bold", "font-size" = "12px"),
      textsize  = "12px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    colors   = legend_colors,
    labels   = legend_labels,
    title    = "Increase in Uncomp. Care",
    position = "bottomright",
    opacity  = 0.85
  )
```

Row
-------------------------------------

### Implications

```{r}
HTML('
<div style="padding:16px 24px; font-family:\"Source Sans Pro\", sans-serif;
            color:#1D1D20; font-size:1.1rem; line-height:1.8;">
  <p>
    Based on this figure, it is evident that eliminating automatic re-enrollment is projected
    to raise the uninsured rate in every state without exception — a decrease is not possible
    under this provision, as no state gains enrollees from the elimination of automatic
    re-enrollment. Notably, the largest increases are concentrated in the South and critically,
    are states that have not expanded Medicaid, meaning their safety-net infrastructure is
    already strained and residents have fewer coverage alternatives when Marketplace enrollment
    lapses further. When looking at the state-level fiscal burden, several states face
    uncompensated care increases in the hundreds of millions to billions of dollars on their own,
    with Florida projected to absorb the largest single-state increase. These disparities carry
    significant equity implications: the communities projected to lose the most coverage are
    disproportionately low-income, already underserved, and located in states with the least
    robust safety-net systems to absorb the resulting uncompensated care burden.
  </p>
</div>
')
```

Row {data-height=120}
-------------------------------------

###

```{r}
HTML('
<div style="padding:10px 24px; font-family:\"Source Sans Pro\", sans-serif;
            font-size:0.82rem; color:#888; border-top:1px solid #e5e5e5;">
  <i>
    Centers for Medicare &amp; Medicaid Services. (2025). <em>2025 Marketplace Open Enrollment Period Public Use Files.</em> U.S. Department of Health and Human Services. &nbsp;|&nbsp;
    Garthwaite, C., Gross, T., &amp; Notowidigdo, M. J. (2015). <em>Hospitals as insurers of last resort</em> (NBER Working Paper No. 21290). National Bureau of Economic Research. https://www.nber.org/papers/w21290 &nbsp;|&nbsp;
    Kaiser Family Foundation. (2024). <em>State health facts: Total population by insurance status</em> [Data set]. KFF. https://www.kff.org/state-health-policy-data/state-indicator/total-population/ &nbsp;|&nbsp;
    U.S. Bureau of Labor Statistics. (2026). <em>CPI inflation calculator.</em> U.S. Department of Labor. https://www.bls.gov/data/inflation_calculator.htm &nbsp;|&nbsp;
    Author Calculations.
  </i>
</div>
')
```


Uncompensated Care Costs
=====================================

Row {data-height=500}
-------------------------------------

### Predicted Total U.S. Uncompensated Care Cost

```{r}
nat_bar <- data.frame(
  category = c("Current Costs", "Increase in Costs"),
  value    = c(nat_current, nat_change),
  label    = c(
    paste0("$", round(nat_current / 1e9, 2), "B"),
    paste0("$", round(nat_change  / 1e9, 2), "B")
  )
)

plot_ly(
  nat_bar,
  x                = ~"United States",
  y                = ~value,
  type             = "bar",
  color            = ~category,
  colors           = c("Current Costs" = col_current, "Increase in Costs" = col_increase),
  text             = ~label,
  textposition     = "inside",
  insidetextanchor = "middle",
  textfont         = list(color = "white", size = 16, family = "Arial Black"),
  hovertemplate    = "<b>%{fullData.name}</b><br>$%{y:,.0f}<extra></extra>"
) %>%
  layout(
    barmode = "stack",
    title   = list(text = "<b>Predicted Total U.S. Uncompensated Care Cost</b>",
                   font = list(size = 14)),
    xaxis  = list(title = ""),
    yaxis  = list(title = "Uncompensated Care Costs ($)", tickformat = "$,.0s"),
    legend = list(orientation = "v", x = 1.02, y = 0.5),
    margin = list(t = 60, b = 90),
    annotations = list(list(
      x = 0, y = -0.26, xref = "paper", yref = "paper",
      text      = "<i>Source: CMS, 2025; Author Calculations</i>",
      showarrow = FALSE,
      font      = list(size = 10, color = "gray")
    ))
  )
```

### Uninsured Rate Scenario Explorer {data-width=420}

```{r}
HTML(paste0('
<div style="padding:20px; font-family:\"Source Sans Pro\", sans-serif; color:#1D1D20;">

  <p style="font-size:0.95rem; color:#444; line-height:1.65; margin-bottom:18px;">
    Enter a hypothetical national uninsured rate to explore how uncompensated care costs
    would change. The <strong>current rate</strong> is
    <strong>', round(nat_current_rate * 100, 2), '%</strong> and the
    <strong>projected post-policy rate</strong> is
    <strong>', round(nat_post_rate * 100, 2), '%</strong>.
  </p>

  <label for="rateInput" style="font-weight:600; font-size:0.95rem;">
    Hypothetical National Uninsured Rate (%):
  </label><br/>
  <input
    type="number"
    id="rateInput"
    value="', round(nat_post_rate * 100, 2), '"
    min="0" max="50" step="0.1"
    style="margin-top:8px; padding:6px 10px; font-size:1rem; width:130px;
           border:1px solid #ccc; border-radius:4px;"
    oninput="updateCalc()"
  />

  <div id="calcResults" style="margin-top:20px; margin-bottom:20px;"></div>

  <div style="font-weight:600; font-size:0.9rem; margin-bottom:8px; color:#555;">
    Hypothetical Cost Breakdown
  </div>
  <div style="display:flex; align-items:center; gap:16px;">
    <div style="display:flex; flex-direction:column; align-items:center; width:140px;">
      <div style="font-size:0.78rem; color:#555; margin-bottom:4px;" id="hypoLabel">—</div>
      <div id="barContainer" style="width:100%; height:150px; border:1px solid #ddd; border-radius:3px; overflow:hidden; display:flex; flex-direction:column; justify-content:flex-end;">
        <div id="hypoChange" style="background:#7B2D42; width:100%; height:0px;"></div>
        <div id="hypoBase"   style="background:#5E663D; width:100%; height:0px;"></div>
      </div>
      <div style="font-size:0.75rem; color:#555; margin-top:5px;">Hypothetical</div>
    </div>
    <div style="display:flex; flex-direction:column; gap:8px; font-size:0.82rem; color:#555;">
      <div style="display:flex; align-items:center; gap:6px;">
        <div style="width:14px; height:14px; background:#7B2D42; border-radius:2px; flex-shrink:0;"></div>
        <span>Increase in Costs</span>
      </div>
      <div style="display:flex; align-items:center; gap:6px;">
        <div style="width:14px; height:14px; background:#5E663D; border-radius:2px; flex-shrink:0;"></div>
        <span>Current Costs</span>
      </div>
    </div>
  </div>

</div>

<script>
var usPop         = ', us_pop, ';
var currentUncomp = ', round(nat_current, 2), ';
var costPerPerson = 1230.73;

function fmtAbs(x) {
  if (Math.abs(x) >= 1e9) return "$" + (x / 1e9).toFixed(2) + "B";
  return "$" + (x / 1e6).toFixed(1) + "M";
}
function fmtDiff(x) {
  var sign = x >= 0 ? "+" : "\u2212";
  var a = Math.abs(x);
  if (a >= 1e9) return sign + "$" + (a / 1e9).toFixed(2) + "B";
  return sign + "$" + (a / 1e6).toFixed(1) + "M";
}

function updateCalc() {
  var raw = parseFloat(document.getElementById("rateInput").value);
  var el  = document.getElementById("calcResults");
  if (isNaN(raw) || raw < 0 || raw > 50) {
    el.innerHTML = "<span style=\'color:#c0392b; font-size:0.9rem;\'>Please enter a value between 0 and 50.</span>";
    return;
  }
  var hypoRate      = raw / 100;
  var hypoUninsured = Math.round(usPop * hypoRate);
  var hypoUncomp    = usPop * hypoRate * costPerPerson;
  var diff          = hypoUncomp - currentUncomp;
  var pct           = (diff / currentUncomp) * 100;
  var col = diff >= 0 ? "#7B2D42" : "#5E663D";

  var rows = [
    ["Hypothetical uninsured rate",  raw.toFixed(2) + "%",                        "#1D1D20"],
    ["Projected uninsured (U.S.)",   hypoUninsured.toLocaleString(),               "#1D1D20"],
    ["Projected uncomp. care costs", fmtAbs(hypoUncomp),                           "#1D1D20"],
    ["Change vs. current costs",     fmtDiff(diff),                                 col],
    ["% change in uncomp. spending", (pct >= 0 ? "+" : "") + pct.toFixed(1) + "%", col]
  ];

  var html = "<table style=\'width:100%; border-collapse:collapse; font-size:0.9rem;\'>";
  rows.forEach(function(r, i) {
    var border = i === 3 ? " border-top:1px solid #ddd;" : "";
    html += "<tr><td style=\'padding:5px 4px; color:#666;" + border + "\'>" + r[0] + "</td>" +
            "<td style=\'padding:5px 4px; font-weight:700; text-align:right; color:" + r[2] + ";" + border + "\'>" + r[1] + "</td></tr>";
  });
  html += "</table>";
  el.innerHTML = html;

  var totalBar    = Math.max(currentUncomp * 1.5, hypoUncomp);
  var barH        = 150;
  var basePx      = Math.round((Math.min(currentUncomp, hypoUncomp) / totalBar) * barH);
  var changePx    = Math.round((Math.abs(diff) / totalBar) * barH);
  var changeColor = diff >= 0 ? "#7B2D42" : "#BEE7B8";

  document.getElementById("hypoLabel").textContent         = fmtAbs(hypoUncomp);
  document.getElementById("hypoBase").style.height         = basePx + "px";
  document.getElementById("hypoChange").style.height       = changePx + "px";
  document.getElementById("hypoChange").style.background   = changeColor;
}
updateCalc();
</script>
'))
```

Row
-------------------------------------

### Implications

```{r}
HTML('
<div style="padding:16px 24px; font-family:\"Source Sans Pro\", sans-serif;
            color:#1D1D20; font-size:1.1rem; line-height:1.8;">
  <p>
    The elimination of automatic re-enrollment for APTC recipients is projected to increase
    total U.S. uncompensated care costs by $6.10 billion annually — a 17.6% rise above the
    current national burden of $34.56 billion (Garthwaite et al., 2015; U.S. Bureau of Labor
    Statistics, 2026; Author Calculations). Because uncompensated care scales directly with
    the number of uninsured individuals seeking care without coverage, a policy that raises
    the uninsured rate by even one percentage point translates into billions of dollars in
    additional costs absorbed by hospitals, safety-net providers, and ultimately, insured
    consumers through higher premiums. Use the scenario explorer to the right to see how
    projected uncompensated care costs shift under different hypothetical uninsured rates —
    the steep dollar changes across even small rate adjustments illustrate just how
    consequential coverage loss is at a population level.
  </p>
</div>
')
```
